{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-07 thesis/education.tex
very subtle
add mutex test to submissions table

\begin{enumerate}
	\item How many Landslide bug reports did you receive from course staff? (Please indicate a number.)
	\item Among those bug reports, how many were you able to diagnose and recognize the root cause in your code? (You may answer ``all'', ``none'', or a number.)
	\item Among those bug reports, how many described a behaviour that you believe was NOT a bug? For example, Landslide printed an execution trace that was actually impossible, or Landslide reported a bug about some behaviour that was actually allowed by the Pintos specification. (You may answer ``all'', ``none'', or a number.)
	\item About how much time did you spend interpreting Landslide's debugging output? (Please indicate a number of minutes, or a range if uncertain, e.g. ``30-60 minutes''.)
	\item I found Landslide's debugging output easy to understand.
		(Multiple choice from strongly disagree to strongly agree.)
	\item It's easier to diagnose the root cause of a bug with Landslide than with a stress test (for example {\tt exec-multiple}).
		(Multiple choice from strongly disagree to strongly agree; plus ``Not sure'' and ``Easier for some bugs but harder for others'')
	\item I feel that by interpreting Landslide's debugging output I learned to understand concurrency better.
		(Multiple choice from strongly disagree to strongly agree.)
	\item These kinds of concurrency bugs are important to fix, even though they don't count against my grade.
		(Multiple choice from strongly disagree to strongly agree.)
	\item Suppose after you submitted your pintos, we gave you 100 CPU-hours on the cloud provider of your choice to test it. Then we extended the project deadline by a day for you to use the results to fix bugs and get partial credit. How would you divide that CPU time between the staff-provided stress tests and Landslide?
		(Multiple choice: 0/10/.../100 CPU-hours on Landslide, 100/90/.../0 CPU-hours on stress tests.)
	\item If course staff were to allow students to resubmit updated code after reviewing Landslide bug reports to receive partial credit for each bug that had been fixed, it would be worth my time to try that (even if I could be spending that time working on the next project instead).
		(Multiple choice from strongly disagree to strongly agree.)
	\item If a friend of mine took OS next semester, I would recommend that they should sign up to receive Landslide bug reports during projects in the future.
		(Multiple choice from strongly disagree to strongly agree.)
	\item Regarding the previous question, why or why not?
	\item Your name
	\item Your project partner's name (if applicable)
\end{enumerate}

@@

\begin{table}[h]
	\begin{center}
		% TODO: calc p values for this
		% note, smth like fisher's method which "combines independent studies into a meta p value"
		% cannot be applied for 2 reasons - 1, these are not independent since the same studence were
		% running landslide on diff test cases; 2, the hypotheses tested are different for each test
	\begin{tabular}{r||cc|cc|c}
		& \multicolumn{2}{c|}{Landslide users} & \multicolumn{2}{c|}{Non-users} & \\
		Bug name & \#correct & total & \#correct & total & $\Delta P(\mathsf{correct})$ \\
		\hline
		%Exit UAF  & 9	& 21	& 2	& 16	& $+30.4\%$ \\ % not filtering out did not run tej; also using "old" thing..?
		Exit UAF  & 11	& 21	& 0	& 16	& $+52.4\%$ \\
		% nb. this is not filter out the did not run paras -- smae for all of the above/below
		% Paraguay  & 22	& 24	& 12	& 13	& $-0.64\%$ \\ % not filtering out did not run para
		Paraguay  & 20	& 21	& 14	& 16	& $+07.7\%$ \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		%Downgrade & 17	& 24	& 8	& 12	& $+4.17\%$ \\ % not filtering out either did not run, or unrel ding
		Downgrade & 10	& 13	& 19	& 23	& $-05.7\%$ \\
		Total & \\ % TODO
	\end{tabular}
	\end{center}
	\caption{
		Correlation of student Landslide use with solving certain bugs in their final submission during Fall 2017 and Spring 2018 semesters.
		Students who used Landslide in general but whose usage snapshots showed they did not run the test case in question are counted as non-Landslide-users here.
		$\Delta P$ stands for the increased probability that a student submits a correct P2
		(with respect to each bug)
		given that they use Landslide, versus not using Landslide; i.e.,
		$P(\mathsf{correct}|\mathsf{landslide})-
		P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

{\bf Paraguay.}
Named after Ivan Jager, the Paraguayan 15-410 TA from 2004-2006 who originally discovered it,
this refers to a condition-variable bug in which a thread which sequentially sleeps on two different condition variables
can spuriously wake up early from the second {\tt cond\_wait()}.
The precise reasoning why many common na\"ive student implementations are susceptible to this bug,
as well as the 3 major ways of fixing it,
are one of 15-410 staff's closely-guarded secrets (see \sect{\ref{sec:410-secrecy}}).
Suffice it to say that the test {\tt paraguay} invokes a specially-made Pathos {\tt misbehave} mode
which biases thread scheduling towards the specific interleaving required to exhibit the bug
(Landslide, of course, replaces this {\tt misbehav}iour with model checking).
% TODO: i feel tihs sentence should go somewhere else, like after analysis, not in description
Hence, despite being a subtle concurrency bug,
the official course test suite is likely to expose it,
so comparing how many Landslide users and non-users submitted this bug in particular
speaks more to the impact of Landslide's preemption traces
in helping students diagnose a bug that a ``stress'' test could already find.
%
{\tt paraguay} is itself the Landslide-friendly test for its eponymous bug.
%, and
%the official rubric penalty for it is
%{\tt :paraguay -2}.
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.

>>>>

\begin{enumerate}
	\item How many Landslide bug reports did you receive from course staff? (Please indicate a number.)
	\item Among those bug reports, how many were you able to diagnose and recognize the root cause in your code? (You may answer ``all'', ``none'', or a number.)
	\item Among those bug reports, how many described a behaviour that you believe was NOT a bug? For example, Landslide printed an execution trace that was actually impossible, or Landslide reported a bug about some behaviour that was actually allowed by the Pintos specification. (You may answer ``all'', ``none'', or a number.)
	\item About how much time did you spend interpreting Landslide's debugging output? (Please indicate a number of minutes, or a range if uncertain, e.g. ``30-60 minutes''.)
	\item I found Landslide's debugging output easy to understand.
		(Multiple choice from strongly disagree to strongly agree.)
	\item It's easier to diagnose the root cause of a bug with Landslide than with a stress test (for example {\tt exec-multiple}).
		(Multiple choice from strongly disagree to strongly agree; plus ``Not sure'' and ``Easier for some bugs but harder for others'')
	\item I feel that by interpreting Landslide's debugging output I learned to understand concurrency better.
		(Multiple choice from strongly disagree to strongly agree.)
	\item These kinds of concurrency bugs are important to fix, even though they don't count against my grade.
		(Multiple choice from strongly disagree to strongly agree.)
	\item Suppose after you submitted your pintos, we gave you 100 CPU-hours on the cloud provider of your choice to test it. Then we extended the project deadline by a day for you to use the results to fix bugs and get partial credit. How would you divide that CPU time between the staff-provided stress tests and Landslide?
		(Multiple choice: 0/10/.../100 CPU-hours on Landslide, 100/90/.../0 CPU-hours on stress tests.)
	\item If course staff were to allow students to resubmit updated code after reviewing Landslide bug reports to receive partial credit for each bug that had been fixed, it would be worth my time to try that (even if I could be spending that time working on the next project instead).
		(Multiple choice from strongly disagree to strongly agree.)
	\item If a friend of mine took OS next semester, I would recommend that they should sign up to receive Landslide bug reports during projects in the future.
		(Multiple choice from strongly disagree to strongly agree.)
	\item Regarding the previous question, why or why not?
	\item Your name.
	\item Your project partner's name (if applicable)
\end{enumerate}

@@

\begin{table}[h]
	\begin{center}
		% TODO: calc p values for this
		% note, smth like fisher's method which "combines independent studies into a meta p value"
		% cannot be applied for 2 reasons - 1, these are not independent since the same studence were
		% running landslide on diff test cases; 2, the hypotheses tested are different for each test
	\begin{tabular}{r||cc|cc|c}
		& \multicolumn{2}{c|}{Landslide users} & \multicolumn{2}{c|}{Non-users} & \\
		Bug name & \#correct & total & \#correct & total & $\Delta P(\mathsf{correct})$ \\
		\hline
		%Exit UAF  & 9	& 21	& 2	& 16	& $+30.4\%$ \\ % not filtering out did not run tej; also using "old" thing..?
		Exit UAF  & 11	& 21	& 0	& 16	& $+52.4\%$ \\
		% nb. this is not filter out the did not run paras -- smae for all of the above/below
		% Paraguay  & 22	& 24	& 12	& 13	& $-0.64\%$ \\ % not filtering out did not run para
		Paraguay  & 20	& 21	& 14	& 16	& $+07.7\%$ \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		%Downgrade & 17	& 24	& 8	& 12	& $+4.17\%$ \\ % not filtering out either did not run, or unrel ding
		Downgrade & 10	& 13	& 19	& 23	& $-05.7\%$ \\ % TODO verify these buggers by eye
		% I spent more hours than I should have looking at this landslide-user-submitted-bug group.
		% Can't name the group here, but if you're me and have the analysis file,
		% it's the one from f17 with 3 mutex_test logs in their "u" snapshot.
		% It seems like a mysterious older version of mutex.c was present between their "a" version
		% (which gets fully verified very quickly btw) and their "u" version (for which landslide finds
		% the bug in just a few minutes, no way they could be running the same version; plus the reported
		% data races and so on in the id-log don't line up to the "u" mutex.c source code).
		% That version was correct, then they updated it (introducing the bug) and ran a bunch of other tests,
		% but didn't rerun mutex_test to check their update.
		Mutex     & 17	& 18	& 19	& 19	& $-05.6\%$ \\
		\hline
		% TODO: figure out if calculating a p value (or w/e) for this the same way is valid,
		% since they summands are arguably not independent? ..or maybe they are?
		Total     & 58	& 73	& 52	& 74	& $+09.2\%$ \\
	\end{tabular}
	\end{center}
	\caption{
		Correlation of student Landslide use with solving certain bugs in their final submission during Fall 2017 and Spring 2018 semesters.
		Students who used Landslide in general but whose usage snapshots showed they did not run the test case in question are counted as non-Landslide-users here.
		$\Delta P$ stands for the increased probability that a student submits a correct P2
		(with respect to each bug)
		given that they use Landslide, versus not using Landslide; i.e.,
		$P(\mathsf{correct}|\mathsf{landslide})-
		P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		Note that the bottom row's totals double-count students,
		which only makes sense if you believe the incidences of each bug within a given submission are independent.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

{\bf Paraguay.}
Named after Ivan Jager, the Paraguayan 15-410 TA from 2004-2006 who originally discovered it,
this refers to a condition-variable bug in which a thread which sequentially sleeps on two different condition variables
can spuriously wake up early from the second {\tt cond\_wait()}.
The precise reasoning why many na\"ive student implementations are susceptible to this bug,
as well as the 3 major ways of fixing it,
are one of 15-410 staff's closely-guarded secrets (\sect{\ref{sec:410-secrecy}}).
Suffice it to say that the {\tt paraguay} test invokes a custom Pathos {\tt misbehave} mode
which biases thread scheduling towards the interleaving required to exhibit the bug
(Landslide, of course, replaces this {\tt misbehav}iour with model checking).
% TODO: i feel tihs sentence should go somewhere else, like after analysis, not in description
Hence, despite being a subtle concurrency bug,
the official course test suite is likely to expose it,
so comparing how many Landslide users and non-users submitted this bug in particular
speaks more to the impact of Landslide's preemption traces
in helping students diagnose a bug that a ``stress'' test could already find.
%
{\tt paraguay} is itself the Landslide-friendly test for its eponymous bug.
%, and
%the official rubric penalty for it is
%{\tt :paraguay -2}.
Whether each student's submitted P2 solved or suffered from this bug
was determined by me personally inspecting their code.
